%Homework2
%Emily Lukshin
%MATH 4255 Spring 2020
%February 17 2020

%Preamble

\documentclass{report}
\usepackage[left=1in,right=1in,top=1in,bottom=1in]{geometry}
\setlength{\headheight}{23pt}
\usepackage{amsthm}
\usepackage{amsfonts}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{mathrsfs}
\usepackage{multirow}
\usepackage{verbatim}
\usepackage{url}
\usepackage{yfonts}
\usepackage{fancyhdr}
\pagestyle{fancy}
\usepackage[colorlinks]{hyperref}
\usepackage{wrapfig}
\usepackage{graphicx}
\begin{document}

%Header

\fancyhead{}
\fancyfoot{}
\lhead{\large \textbf{Math 4255/5255}}
\chead{\Large \textbf{Homework 2}}
\rhead{\large \textbf{Emily Lukshin}}

%Custom Commands

\newcommand{\N}{\mathbb{N}}
\newcommand{\Z}{\mathbb{Z}}
\newcommand{\Q}{\mathbb{Q}}
\newcommand{\R}{\mathbb{R}}
\newcommand{\C}{\mathbb{C}}
\newcommand{\D}{\mathbb{D}}
\newcommand{\<}{\left\langle}
\renewcommand{\>}{\right\rangle}
\renewcommand{\Re}[1]{\text{Re}\ #1}
\renewcommand{\Im}[1]{\text{Im}\ #1}
\renewcommand{\mod}[1]{(\operatorname{mod}#1)}

%Document

\noindent
{\bf Please submit your solutions (both tex and PDF files) to WyoCourses; and hand in a printed version in class. If you are a graduate student, please also finish Problem 5 and 6.}

\medskip

The following problems are mainly selected from Ross's book. 

\begin{enumerate}

%EXERCISE 1

\item Let $E$ and $F$ be two events. Prove that
\[
  {P}(E\cap F^c)=P(E)-P(E\cap F). 
\]

{\bf Solution:}\\
    $E\cap F$ and $E\cap F^c$ are mutually exclusive, therefore
    $$P((E\cap F) \cup (E \cap F^c))=P(E\cap F)+P(E\cap F^c)$$
    We know
    $$(E\cap F)\cup(E\cap F^c)=E(F\cup F^c)\quad \text{By the distributive laws}$$
    $$=E(1)$$
    $$=E$$
    Therefore,
    $$P(E)=P((E\cap F)\cup(E\cap F^c))$$
    $$=P(E\cap F)+P(E\cap F^c)$$
    By rearranging, we get
    $$P(E\cap F^c)=P(E)-P(E\cap F)$$

%EXERCISE 2

\item Suppose that $A$ and $B$ are mutually exclusive events for with $P(A)=0.3$ and $P(B)=0.5$. What is the probability that
\begin{enumerate}
	\item either $A$ or $B$ occurs?
	\item $A$ occurs but not $B$?
	\item both $A$ and $B$ occur. 
\end{enumerate}

{\bf Solution:}
\begin{enumerate}
    \item $(A\cup B)=P(A)+P(B)=0.3+0.5=0.8$
    \item $P(A)=0.3$, because $A$ and $B$ are mutually exclusive, the occurrence of $A$ means that there is no occurrence of $B$.
    \item $P(A\cap B)=P(\emptyset)=0$, because $A$ and $B$ are mutually exclusive, $A$ and $B$ cannot occur at the same time.
\end{enumerate}

%EXERCISE 3

\item A pair of fair dice are rolled.
\begin{enumerate}
    \item Describe the outcome of the experiment and the sample space S by using vectors (in this case, a vector has two coordinates).
    \item Compute the number of elements of the sample space S.
    \item What is the probability to get a double 6?
    \item What is the probability that the second die lands on a higher value than does the first?

\end{enumerate}

{\bf Solution:}

\begin{enumerate}
    \item $S=\{(i,j):i,j=1,2,3,4,5,6\}$, where the outcome $(i,j)$ is said to occur if $i$ appears on the leftmost die and $j$ on the other die.
    \item $6^2=36$
    \item $\frac{1}{6^2}=\frac{1}{36}$
    \item $\frac{{6 \choose 2}}{6^2}=\frac{15}{36}=\frac{5}{12}$
\end{enumerate}
	
%EXERCISE 4
	
\item Prove \emph{Bonferroni's inequality}, namely, 
\[
  P(EF)\geq P(E)+P(F)-1.
\]

{\bf Solution:}\\
From Axiom 1 of probability, we know
$$P(E\cup F)\leq 1$$
$$\Longrightarrow P(E)+P(F)-P(E\cap F)\leq 1$$
$$\Longrightarrow P(E)+P(F)-1\leq P(E\cap F)$$
By rearranging, we get
$$P(E\cap F)\geq P(E)+P(F)-1$$
	
\end{enumerate}



\end{document}​\documentclass{article}
\usepackage[utf8]{inputenc}

\title{hw-2}
\author{elukshin }
\date{February 2020}

\begin{document}

\maketitle

\section{Introduction}

\end{document}
